Fetal Brain isoform analysis - junctions
========================================================

Gloria Li         
`r date()` 

<!-- re-knit after modify junction_valid.R or junction.R script -->

```{r include = FALSE}
opts_chunk$set(message=FALSE, echo = FALSE, warning = FALSE)
```

```{r data}
setwd("~/快盘/FetalBrain/RNAseq/junction")
load("junction_valid.Rdata")
library(ggplot2)
```

## Validate previously identified isoforms with junction RPKM
  * Previous isoform identification with DE exons
    * DE exons by DEfine FDR = 0.01
    * Exon RPKM $\ge$ 10% gene RPKM in one sample & $\le$ 1% in the other
    * Gene RPKM of both samples > 0.005 
    * Exclude DE genes by DEfine FDR = 0.01

  * Validation: For each isoform exon in the previous pairwise comparison
    * Find junctions associated with this exon with enough coverage, i.e. sum of junction coverage of two samples $\ge$ 1
    * Identify junctions that RPKM change in the same direction as the exon
    * Junction RPKM > 0.1 in one sample and < 0.1 in the other      
  
## Results: 

