---
title: "FetalBrain - Figures"
author: "Gloria Li"
date: "November 4, 2014"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_depth: 4
---

Updated: `r date()`

```{r setup, include = FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(xtable)
library(VennDiagram)
library(gridExtra)
load("~/快盘/FetalBrain/MeDIP/DMR_MZ.Rdata")
load("~/快盘/FetalBrain/MeDIP/DMR_neurospheres.Rdata")
load("~/快盘/FetalBrain/WGBS/WGBS.DMR.Rdata")
load("~/快盘/FetalBrain/RNAseq/DEfine/gene/FetalBrain_DEgenes.Rdata")
load("~/快盘/FetalBrain/RNAseq/isoform/FetalBrain_isoform.Rdata")
load("~/快盘/FetalBrain/RNAseq/junction/FetalBrain_isoform_valid.Rdata")
load("~/快盘/FetalBrain/RNAseq/epiProfile/exonProfile_Rmd.Rdata")
knitr::opts_chunk$set(message=FALSE, echo = FALSE, warning = FALSE, results = FALSE, fig.height = 8, fig.width = 10)
```

## Figure 1: MZ twins 
### Figure 1a: No. of UMRs between MZ twins and neurospheres 
  * No. of UMRs (hypermethylated in red and hypomethylated in blue) between monozygotic twins and Cortex vs GE neurospheres.    
  * No. of UMRs between MZ twins are on the same scale as UMRs between neurospheres.   
  
```{r MZ_scale}
MZ_scale <- data.frame(Sample = rep(c(DMR_MZ_summary$Sample, DMR_neurospheres_summary$Sample), 5), 
                        Category = rep(c(rep("MZ twins", nrow(DMR_MZ_summary)), rep("Neurospheres", nrow(DMR_neurospheres_summary))), 5), 
                        DM_DE = c(rep(c("Hyper / UP", "Hypo / DN"), each = nrow(DMR_MZ_summary) + nrow(DMR_neurospheres_summary)), rep(c("Hyper / UP", "Hypo / DN"), each = nrow(DMR_MZ_summary) + nrow(DMR_neurospheres_summary)), rep("Isoforms", nrow(DMR_MZ_summary) + nrow(DMR_neurospheres_summary))), 
                        Assay = factor(c(rep(c("UMRs", "DE genes"), each = (nrow(DMR_MZ_summary) + nrow(DMR_neurospheres_summary))*2), rep("Isoforms", (nrow(DMR_MZ_summary) + nrow(DMR_neurospheres_summary)))), levels = c("UMRs", "DE genes", "Isoforms")), 
                        value = c(DMR_MZ_summary$Hyper.DMR, DMR_neurospheres_summary$Hyper.DMR, -DMR_MZ_summary$Hypo.DMR, -DMR_neurospheres_summary$Hypo.DMR, individual_DE_summary$UP[1:3], cortex_GE_DE_summary$UP[1:2], -individual_DE_summary$DN[1:3], -cortex_GE_DE_summary$DN[1:2], individual_summary[1:3, "isoform_genes"], cortex_GE_summary[1:2, "isoform_genes"]))
MZ_scale$Sample <- factor(gsub("-HuFNSC", "", MZ_scale$Sample), levels = c("Brain01_Brain02", "Cortex01_Cortex02", "GE01_GE02", "Cortex01_GE01", "Cortex02_GE02"))
(MZ_scale_figure <- ggplot(MZ_scale, aes(x = Sample, y = value, fill = DM_DE)) + 
   geom_bar(stat="identity", position = "identity", width = 0.5) + 
   geom_hline(yintercept = 0) + 
   facet_grid(Assay ~ Category, scales = "free", space = "free_x") + 
   scale_fill_manual(values = c("red", "blue", "black"), name = "") + 
   scale_y_continuous(breaks = c(seq(-3000, 3000, by = 1000), -500, 500), labels = c(abs(seq(-3000, 3000, by = 1000)), 500, 500)) + 
   xlab("") + 
   ylab("No. of UMRs") + 
   theme_bw() + 
   theme(strip.text = element_text(size = 15), axis.title.y = element_text(size = 15), axis.text = element_text(size = 12), legend.text = element_text(size = 15)))
ggsave(MZ_scale_figure, file = "~/快盘/FetalBrain/Figures/Figure1a.pdf")
```

### Figure 1b: Asymmetry in UMRs between MZ twins in Brain and Cortex
  * UMR frequency (base pair per million base pair) between MZ twins (HuFNSC01 UMRs in blue, and HuFNSC02 UMRs in red) in Brain, Neurospheres Cortex derived and GE derived across different chormosomes. 
  * HuFNSC02 UMRs are __`r round(sum(DMR_freq_MZ_figure$data[DMR_freq_MZ_figure$data$cell == "Brain" & DMR_freq_MZ_figure$data$DM == "1", "freq"])/-sum(DMR_freq_MZ_figure$data[DMR_freq_MZ_figure$data$cell == "Brain" & DMR_freq_MZ_figure$data$DM == "-1", "freq"]), 2)`__-fold enriched in Brain, and __`r round(sum(DMR_freq_MZ_figure$data[DMR_freq_MZ_figure$data$cell == "Cortex" & DMR_freq_MZ_figure$data$DM == "1", "freq"])/-sum(DMR_freq_MZ_figure$data[DMR_freq_MZ_figure$data$cell == "Cortex" & DMR_freq_MZ_figure$data$DM == "-1", "freq"]), 2)`__-fold enriched in Cortex compared to HuFNSC01 UMRs.    
  
```{r MZ_asymmetry}
(DMR_freq_MZ_figure <- DMR_freq_MZ_figure + ylab("UMR frequency (bp/MB)") + xlab(""))
ggsave(DMR_freq_MZ_figure, file = "~/快盘/FetalBrain/Figures/Figure1b.pdf")
```

### Figure 1c: Differential expressed genes between MZ twins are cell type specific
  * No. of differential expressed genes between MZ twins in Brain (green), Neurospheres Cortex derived (red), and GE derived (blue).  
  
```{r MZ_DE, fig.height=5}
MZ_UP <- list(Brain = as.character(filter(brain01_brain02DE, DE == "UP")[, "V1"]), Cortex = as.character(filter(cortex01_cortex02DE, DE == "UP")[, "V1"]), GE = as.character(filter(GE01_GE02DE, DE == "UP")[, "V1"]))
venn_MZ_UP <- venn.diagram(MZ_UP, filename = NULL, fill = c("green", "red", "blue"), main = "UP", main.cex = 1.5)
MZ_DN <- list(Brain = as.character(filter(brain01_brain02DE, DE == "DN")[, "V1"]), Cortex = as.character(filter(cortex01_cortex02DE, DE == "DN")[, "V1"]), GE = as.character(filter(GE01_GE02DE, DE == "DN")[, "V1"]))
venn_MZ_DN <- venn.diagram(MZ_DN, filename = NULL, fill = c("green", "red", "blue"), main = "DN", main.cex = 1.5)
grid.arrange(gTree(children = venn_MZ_UP), gTree(children = venn_MZ_DN), nrow = 1)
pdf("~/快盘/FetalBrain/Figures/Figure1c.pdf", height = 5, width = 10)
grid.arrange(gTree(children = venn_MZ_UP), gTree(children = venn_MZ_DN), nrow = 1)
dev.off()
```

## Figure 2: Neurospheres Cortex and GE
### Figure 2a: UMR/DE genes accumulates over time

```{r neurospheres_scale}
neurospheres_scale <- data.frame(Donor = c(rep(c("HuFNSC02", "HuFNSC04", "HuFNSC01", "HuFNSC02", "HuFNSC03", "HuFNSC04"), 2), "HuFNSC01", "HuFNSC02", "HuFNSC03", "HuFNSC04"), 
                                 GW = c(rep(c("17-week", "13-week", "17-week", "17-week", "13-week", "13-week"), 2), "17-week", "17-week", "13-week", "13-week"), 
                                 Assay = factor(c(rep(c(rep("UMRs", 2), rep("DE genes", 4)), 2), rep("Isoforms", 4)), levels = c("UMRs", "DE genes", "Isoforms")), 
                                 DM_DE = c(rep(c("Cortex UMR/UP", "GE UMR/UP"), each = 6), rep("Isoform", 4)), 
                                 value = c(DMR_WGBS_summary$Hypo.DMR, cortex_GE_DE_summary$UP, DMR_WGBS_summary$Hyper.DMR, cortex_GE_DE_summary$DN, cortex_GE_summary[, "isoform_genes"]))
neurospheres_scale[neurospheres_scale$DM_DE == "GE UMR/UP", "value"] <- -neurospheres_scale[neurospheres_scale$DM_DE == "GE UMR/UP", "value"]
(neurospheres_scale_figure <- ggplot(neurospheres_scale, aes(x = Donor, y = value, fill = DM_DE)) + 
   geom_bar(stat="identity", position = "identity", width = 0.4) + 
   geom_hline(yintercept = 0) + 
   facet_grid(Assay ~ GW, scales = "free") + 
   scale_fill_manual(values = c("red", "blue", "black"), name = "") + 
   #scale_y_continuous(breaks = seq(-3000, 3000, by = 1000), labels = abs(seq(-3000, 3000, by = 1000))) + 
   xlab("") + 
   ylab("No. of UMRs / genes") + 
   theme_bw() + 
   theme(strip.text = element_text(size = 15), axis.title.y = element_text(size = 15), axis.text = element_text(size = 12), legend.text = element_text(size = 15)))
ggsave(neurospheres_scale_figure, file = "~/快盘/FetalBrain/Figures/Figure2a.pdf")
```

### Figure 2b: UMR asymmetry between Neurospheres Cortex and GE
  * UMR frequency (base pair per million base pair) between neurospheres cortex derived (red), and GE derived(blue). 
  * On average, there are , __`r round(sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$DM == "-1", "freq"])/-sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$DM == "1", "freq"]), 2)`__-fold enrichment in total UMR frequency in Cortex compared to GE, __`r round(sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$donor == "HuFNSC02" & DMR_freq_WGBS_figure$data$DM == "-1", "freq"])/-sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$donor == "HuFNSC02" & DMR_freq_WGBS_figure$data$DM == "1", "freq"]), 2)`__ in HuFNSC02, and __`r round(sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$donor == "HuFNSC04" & DMR_freq_WGBS_figure$data$DM == "-1", "freq"])/-sum(DMR_freq_WGBS_figure$data[DMR_freq_WGBS_figure$data$donor == "HuFNSC04" & DMR_freq_WGBS_figure$data$DM == "1", "freq"]), 2)`__ in HuFNSC04.    

```{r neurospheres_asymmetry}
(DMR_freq_WGBS_figure <- DMR_freq_WGBS_figure + scale_fill_manual(values = c("blue", "red"), labels = c("GE UMRs", "Cortex UMRs"), name = ""))
ggsave(DMR_freq_WGBS_figure, file = "~/快盘/FetalBrain/Figures/Figure2b.pdf")
```

### Figure 2c: GREAT enrichment for neurospheres UMRs

```{r neurospheres_GREAT}
GOBP_Cortex02.UMR_WGBS <- read.delim("~/快盘/FetalBrain/WGBS/DMR/enrich/ALL_GREAT_GOBP_DMR.Cortex-HuFNSC02_GE-HuFNSC02.hypo.tsv", head = F, as.is = T, skip = 4)
GOBP_GE02.UMR_WGBS <- read.delim("~/快盘/FetalBrain/WGBS/DMR/enrich/ALL_GREAT_GOBP_DMR.Cortex-HuFNSC02_GE-HuFNSC02.hyper.tsv", head = F, as.is = T, skip = 4)
GOBP_Cortex02_GE02_WGBS <- rbind(select(mutate(filter(GOBP_Cortex02.UMR_WGBS, V8 >= 2 & V7 <= 0.05 & V16 <= 0.05), UMR = "Cortex.UMR", FDR = -log10(V7), Term = V3), UMR, Term, FDR),
                                 select(mutate(filter(GOBP_GE02.UMR_WGBS, V8 >= 2 & V7 <= 0.05 & V16 <= 0.05), UMR = "GE.UMR", FDR = log10(V7), Term = V3), UMR, Term, FDR))
Terms <- unique(c(arrange(filter(GOBP_Cortex02_GE02_WGBS, UMR == "Cortex.UMR"), desc(FDR))[1:20, "Term"], rev(arrange(filter(GOBP_Cortex02_GE02_WGBS, UMR == "GE.UMR"), FDR)[1:20, "Term"])))
GOBP_Cortex02_GE02_WGBS <- mutate(filter(GOBP_Cortex02_GE02_WGBS, Term %in% Terms), Term = factor(Term, levels = rev(Terms)))
(GOBP_Cortex02_GE02_WGBS_figure <- ggplot(data = GOBP_Cortex02_GE02_WGBS, aes(Term, FDR, fill = UMR)) +
  geom_bar(position = "identity", stat = "identity", width = .5) + 
  coord_flip() + 
  xlab("") + 
  ylab("-log10(FDR)") + 
  scale_fill_manual(values = c("blue", "red"), name = "") + 
  theme_bw())
ggsave(GOBP_Cortex02_GE02_WGBS_figure, file = "~/快盘/FetalBrain/Figures/Figure2c.pdf")
```

### Figure 2d: TFBS asymmetry between neurosphere UMRs

```{r neurospheres_TFBS}
(DMR_TF_figure <- DMR_TF_figure + xlab("Abundance near Cortex UMRs") + ylab("Abundance near GE UMRs")) 
ggsave(DMR_TF_figure, file = "~/快盘/FetalBrain/Figures/Figure2d.pdf")
```

### Figure 2e: DNA methylation at exon boundaries

```{r neurospheres_epiProfile_5mC}
gt <- ggplot_gtable(ggplot_build(WGBS_figure_HuFNSC02 + ggtitle("WGBS HuFNSC02 cortex vs GE"))) 
gt$heights[[4]] <- unit(3.5, "null") 
grid.newpage()
grid.draw(gt) 
grid.text("Average DNA methylation", x = unit(0.015, "npc"), y = unit(0.65, "npc"), rot = 90)
grid.text("CpG density", x = unit(0.015, "npc"), y = unit(0.15, "npc"), rot = 90)

gt <- ggplot_gtable(ggplot_build(WGBS_figure_HuFNSC04 + ggtitle("WGBS HuFNSC04 cortex vs GE"))) 
gt$heights[[4]] <- unit(3.5, "null") 
grid.newpage()
grid.draw(gt) 
grid.text("Average DNA methylation", x = unit(0.015, "npc"), y = unit(0.65, "npc"), rot = 90)
grid.text("CpG density", x = unit(0.015, "npc"), y = unit(0.15, "npc"), rot = 90)



pdf("~/快盘/FetalBrain/Figures/Figure2e.pdf")
grid.draw(gt) 
grid.text("Average DNA methylation", x = unit(0.015, "npc"), y = unit(0.65, "npc"), rot = 90)
grid.text("CpG density", x = unit(0.015, "npc"), y = unit(0.15, "npc"), rot = 90)
dev.off()
```

## Supplemental Figures
### Supplemental 1: Location of UMRs across the genome 

```{r UMR_pos}
(DMR_pos_MZ_ggplot <- DMR_pos_MZ_figure@ggplot + theme(legend.title = element_text(size = 0), panel.margin = unit(0.1, "lines")) + ggtitle("UMRs between MZ twins") + xlab("Position of UMRs on the chromosome") + scale_color_manual(values = c("red", "blue"), labels = c("HuFNSC02 UMRs", "HuFNSC01 UMRs"), name = ""))
(DMR_pos_neurospheres_ggplot <- DMR_pos_WGBS_figure@ggplot + theme(legend.title = element_text(size = 0), panel.margin = unit(0.1, "lines")) + ggtitle("UMRs between neurospheres") + xlab("Position of UMRs on the chromosome") + scale_color_manual(values = c("blue", "red"), labels = c("GE UMRs", "Cortex UMRs"), name = ""))
pdf("~/快盘/FetalBrain/Figures/FigureS1.pdf", height = 10, width = 20)
grid.arrange(DMR_pos_MZ_ggplot, DMR_pos_neurospheres_ggplot, nrow = 1, widths = 3:2)
dev.off()
```

### Supplemental 2: Genomic breakdown of UMRs

```{r genomicBreak}
genomicBreak_MZ_tall <- genomicBreak_MZ_figure$data %>% mutate(DM = revalue(factor(DM), c("hyper" = "HuFNSC02 UMRs", "hypo" = "HuFNSC01 UMRs")))
(genomicBreak_MZ_figure <- ggplot(genomicBreak_MZ_tall, aes(x = Region, y = CpG, fill = Sample)) + 
  geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
  xlab("") + 
  ylab("Fraction of CpG") + 
  ggtitle("UMRs between MZ twins") + 
  facet_wrap(~ DM) + 
  coord_flip() + 
  scale_fill_manual(values = c("green", "red", "blue"), labels = c("Brain", "Cortex", "GE"), name = "") + 
  theme_bw())
genomicBreak_WGBS_tall <- genomicBreak_WGBS_figure$data %>% mutate(DM = revalue(factor(DM), c("hyper" = "GE UMRs", "hypo" = "Cortex UMRs")))
(genomicBreak_WGBS_figure <- ggplot(genomicBreak_WGBS_tall, aes(x = Region, y = CpG, fill = Sample)) + 
  geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
  xlab("") + 
  ylab("Fraction of CpG") + 
  ggtitle("UMRs between neurospheres") + 
  coord_flip() + 
  scale_fill_manual(values = c("red", "blue"), labels = c("HuFNSC02", "HuFNSC04"), name = "") + 
  facet_wrap(~ DM) + 
  theme_bw())
pdf("~/快盘/FetalBrain/Figures/FigureS2.pdf", height = 7, width = 15)
grid.arrange(genomicBreak_MZ_figure, genomicBreak_WGBS_figure, nrow = 1)
dev.off()
```

### Supplemental 3: Fraction of DE genes with proximal UMRs

```{r proximal_DE, fig.height=5, fig.width=7}
DMR_proximal_DE_MZ_summary <- data.frame(Sample = rownames(DMR_proximal_MZ_summary), Total.UP = rep(c(nrow(filter(brain01_brain02DE, DE == "UP")), nrow(filter(cortex01_cortex02DE, DE == "UP")), nrow(filter(GE01_GE02DE, DE == "UP"))), each = 2), Total.DN = rep(c(nrow(filter(brain01_brain02DE, DE == "DN")), nrow(filter(cortex01_cortex02DE, DE == "DN")), nrow(filter(GE01_GE02DE, DE == "DN"))), each = 2), DMR.DE = DMR_proximal_MZ_summary[, "unique.DE.genes"], same.direction = DMR_proximal_MZ_summary[, "same.direction"], DMR.UP = NA, DMR.DN = NA)
DMR_proximal_DE_MZ_summary$DM <- gsub(".*_", "", DMR_proximal_DE_MZ_summary$Sample)
DMR_proximal_DE_MZ_summary$Sample <- gsub("_hy.*", "", DMR_proximal_DE_MZ_summary$Sample)
DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "DMR.UP"] <- (DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "DMR.DE"] - DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "same.direction"]) / DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "Total.UP"]
DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "DMR.UP"] <- DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "same.direction"] / DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "Total.UP"]
DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "DMR.DN"] <- DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "same.direction"] / DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hyper", "Total.DN"]
DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "DMR.DN"] <- (DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "DMR.DE"] - DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "same.direction"])  / DMR_proximal_DE_MZ_summary[DMR_proximal_DE_MZ_summary$DM == "hypo", "Total.DN"]
DMR_proximal_DE_MZ_summary_tall <- data.frame(Sample = rep(DMR_proximal_DE_MZ_summary$Sample, 2), DM = rep(DMR_proximal_DE_MZ_summary$DM, 2), DE = rep(c("UP", "DN"), each = nrow(DMR_proximal_DE_MZ_summary)), Percent = c(DMR_proximal_DE_MZ_summary$DMR.UP, DMR_proximal_DE_MZ_summary$DMR.DN))
DMR_proximal_DE_MZ_summary_tall <- DMR_proximal_DE_MZ_summary_tall %>% mutate(DM = revalue(factor(DM), c("hyper" = "HuFNSC02 UMRs", "hypo" = "HuFNSC01 UMRs")))
(DMR_proximal_DE_MZ_figure <- ggplot(DMR_proximal_DE_MZ_summary_tall, aes(x = Sample, y = Percent, fill = DE)) + 
   geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
   facet_wrap(~ DM) + 
   coord_flip() + 
   scale_fill_manual(values = c("blue", "red"), labels = c("UP in HuFNSC02", "UP in HuFNSC01"), name = "") + 
   xlab("") + 
   ylab("Fraction of DE genes") + 
   ggtitle("UMRs between MZ twins") + 
   theme_bw())
DMR_proximal_DE_neurospheres_summary <- data.frame(Sample = rownames(DMR_WGBS_gene_summary), Total.UP = rep(c(nrow(filter(cortex02_GE02DE, DE == "UP")), nrow(filter(cortex04_GE04DE, DE == "UP"))), each = 2), Total.DN = rep(c(nrow(filter(cortex02_GE02DE, DE == "DN")), nrow(filter(cortex04_GE04DE, DE == "DN"))), each = 2), DMR.DE = DMR_WGBS_gene_summary[, "unique.DE.Genes"], same.direction = DMR_WGBS_gene_summary[, "same.direction"], DMR.UP = NA, DMR.DN = NA)
DMR_proximal_DE_neurospheres_summary$DM <- gsub("-.*", "", DMR_proximal_DE_neurospheres_summary$Sample)
DMR_proximal_DE_neurospheres_summary$Sample <- gsub(".*-", "", DMR_proximal_DE_neurospheres_summary$Sample)
DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "DMR.UP"] <- (DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "DMR.DE"] - DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "same.direction"]) / DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "Total.UP"]
DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "DMR.UP"] <- DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "same.direction"] / DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "Total.UP"]
DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "DMR.DN"] <- DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "same.direction"] / DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "GE_UMRs", "Total.DN"]
DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "DMR.DN"] <- (DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "DMR.DE"] - DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "same.direction"])  / DMR_proximal_DE_neurospheres_summary[DMR_proximal_DE_neurospheres_summary$DM == "Cortex_UMRs", "Total.DN"]
DMR_proximal_DE_neurospheres_summary_tall <- data.frame(Sample = rep(DMR_proximal_DE_neurospheres_summary$Sample, 2), DM = rep(DMR_proximal_DE_neurospheres_summary$DM, 2), DE = rep(c("UP", "DN"), each = nrow(DMR_proximal_DE_neurospheres_summary)), Percent = c(DMR_proximal_DE_neurospheres_summary$DMR.UP, DMR_proximal_DE_neurospheres_summary$DMR.DN))
(DMR_proximal_DE_neurospheres_figure <- ggplot(DMR_proximal_DE_neurospheres_summary_tall, aes(x = Sample, y = Percent, fill = DE)) + 
   geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
   facet_wrap(~ DM) + 
   coord_flip() + 
   scale_fill_manual(values = c("blue", "red"), labels = c("UP in GE", "UP in Cortex"), name = "") + 
   scale_y_continuous(breaks = c(0.01, 0.02)) + 
   xlab("") + 
   ylab("Fraction of DE genes") + 
   ggtitle("UMRs between neurospheres") + 
   theme_bw())
pdf("~/快盘/FetalBrain/Figures/FigureS3.pdf", height = 5, width = 15)
grid.arrange(DMR_proximal_DE_MZ_figure, DMR_proximal_DE_neurospheres_figure, nrow = 1)
dev.off()
```

### Supplemental 4: TFBS asymmetry in MZ UMRs

```{r MZ_TFBS}
(DMR_MZ_TF_figure <- DMR_MZ_TF_figure + xlab("Abundance near HuFNSC02 UMRs") + ylab("Abundance near HuFNSC01 UMRs")) 
ggsave(DMR_MZ_TF_figure, file = "~/快盘/FetalBrain/Figures/FigureS4.pdf", height = 5, width = 15)
```

### Supplemental 5: DAVID enrichment for MZ DE genes

```{r MZ_DE_DAVID}
DAVID_brain01_brain02DE <- read.delim("~/快盘/FetalBrain/RNAseq/DEfine/gene/enrich/brain01_brain02DE_DAVID.txt", as.is = T) %>% filter(Category == "GOTERM_BP_5")
DAVID_cortex01_cortex02DE <- read.delim("~/快盘/FetalBrain/RNAseq/DEfine/gene/enrich/cortex01_cortex02DE_DAVID.txt", as.is = T) %>% filter(Category == "GOTERM_BP_5")
DAVID_GE01_GE02DE <- read.delim("~/快盘/FetalBrain/RNAseq/DEfine/gene/enrich/GE01_GE02DE_DAVID.txt", as.is = T) %>% filter(Category == "GOTERM_BP_5")
Terms <- unique(c(DAVID_brain01_brain02DE[DAVID_brain01_brain02DE$FDR < 0.01, "Term"], DAVID_cortex01_cortex02DE[DAVID_cortex01_cortex02DE$FDR < 0.01, "Term"], DAVID_GE01_GE02DE[DAVID_GE01_GE02DE$FDR < 0.01, "Term"]))
DAVID_MZ <- rbind(filter(DAVID_brain01_brain02DE, Term %in% Terms) %>% mutate(Cell = "Brain"), filter(DAVID_cortex01_cortex02DE, Term %in% Terms) %>% mutate(Cell = "Cortex"), filter(DAVID_GE01_GE02DE, Term %in% Terms) %>% mutate(Cell = "GE")) %>% mutate(Term = gsub("GO:\\d+~", "", Term), FDR = -log10(FDR) -2) %>% select(Cell, Term, FDR)
DAVID_MZ$Term <- factor(DAVID_MZ$Term, levels = rev(unique(DAVID_MZ[order(DAVID_MZ$FDR, decreasing = T), "Term"])))
(DAVID_MZ_figure <- ggplot(DAVID_MZ, aes(x = Term, y = FDR, fill = Cell)) + 
   geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
   coord_flip() + 
   scale_fill_manual(values = c("green", "red", "blue"), labels = c("Brain", "Cortex", "GE"), name = "") + 
   scale_y_continuous(breaks = seq(-4, 6, by = 2), labels = seq(-2, 8, by = 2)) + 
   xlab("") + 
   ylab("-log10(FDR)") + 
   theme_bw())
ggsave(DAVID_MZ_figure, file = "~/快盘/FetalBrain/Figures/FigureS5.pdf")
```

### Supplemental 6: Validate WGBS UMRs with MeDIP/MRE

```{r WGBS_valid}
(valid_boxplot)
ggsave(valid_boxplot, file = "~/快盘/FetalBrain/Figures/FigureS6.pdf")
```

### Supplemental 7: Neurosphere UMR enrichemnt at chromosome ends

```{r neurospheres_chrEnd}
chrEnd_Cortex02UMR <- read.delim("~/快盘/FetalBrain/WGBS/DMR/DMR.Cortex-HuFNSC02_GE-HuFNSC02.m0.75.p0.005.d0.5.s300.c3.hypo.bed.chrom.window", head = F, as.is = T)
chrEnd_GE02UMR <- read.delim("~/快盘/FetalBrain/WGBS/DMR/DMR.Cortex-HuFNSC02_GE-HuFNSC02.m0.75.p0.005.d0.5.s300.c3.hyper.bed.chrom.window", head = F, as.is = T)
chrEnd_Cortex04UMR <- read.delim("~/快盘/FetalBrain/WGBS/DMR/DMR.Cortex-HuFNSC04_GE-HuFNSC04.m0.75.p0.005.d0.5.s300.c3.hypo.bed.chrom.window", head = F, as.is = T)
chrEnd_GE04UMR <- read.delim("~/快盘/FetalBrain/WGBS/DMR/DMR.Cortex-HuFNSC04_GE-HuFNSC04.m0.75.p0.005.d0.5.s300.c3.hyper.bed.chrom.window", head = F, as.is = T)
chrEnd <- rbind(data.frame(DM = "Cortex.UMRs", Donor = "HuFNSC02", chrEnd_Cortex02UMR %>% group_by(V4) %>% summarise(N = sum(V5))), data.frame(DM = "GE.UMRs", Donor = "HuFNSC02", chrEnd_GE02UMR %>% group_by(V4) %>% summarise(N = sum(V5))), data.frame(DM = "Cortex.UMRs", Donor = "HuFNSC04", chrEnd_Cortex04UMR %>% group_by(V4) %>% summarise(N = sum(V5))), data.frame(DM = "GE.UMRs", Donor = "HuFNSC04", chrEnd_Cortex04UMR %>% group_by(V4) %>% summarise(N = sum(V5))))
(chrEnd_figure <- ggplot(chrEnd, aes(x = Donor, y = N, fill = as.factor(V4))) + 
   geom_bar(stat = "identity", width = 0.5, position = "fill") + 
   facet_grid(DM ~ .) + 
   xlab("") + 
   ylab("Fraction of total UMRs") + 
   coord_flip() + 
   scale_fill_discrete(name = "Normalized length \nalong chromosome", labels = as.character(seq(0.1, 1, by = 0.1))) + 
   theme_bw() + 
   theme(strip.text = element_text(size = 15), axis.title.y = element_text(size = 15), axis.text = element_text(size = 12), legend.text = element_text(size = 15)))
ggsave(chrEnd_figure, file = "~/快盘/FetalBrain/Figures/FigureS7.pdf")
```

### Supplemental 8: UMR asymmetry between Cortex and GE with MeDIP

```{r neurospheres_asymmetry_MeDIP}
(DMR_freq_neurospheres_figure <- DMR_freq_neurospheres_figure + scale_fill_manual(values = c("blue", "red"), labels = c("GE UMRs", "Cortex UMRs"), name = ""))
ggsave(DMR_freq_neurospheres_figure, file = "~/快盘/FetalBrain/Figures/FigureS8.pdf")
```

### Supplemental 9: GREAT enrichment for neurospheres UMRs with MeDIP

```{r neurospheres_GREAT_MeDIP}
GOBP_Cortex02.UMR_MeDIP <- read.delim("~/快盘/FetalBrain/MeDIP/DMR/enrich/ALL_GREAT_GOBP_DMR.Cortex-HuFNSC02_GE-HuFNSC02.hypo.tsv", head = F, as.is = T, skip = 4)
GOBP_GE02.UMR_MeDIP <- read.delim("~/快盘/FetalBrain/MeDIP/DMR/enrich/ALL_GREAT_GOBP_DMR.Cortex-HuFNSC02_GE-HuFNSC02.hyper.tsv", head = F, as.is = T, skip = 4)
GOBP_Cortex02_GE02_MeDIP <- rbind(select(mutate(filter(GOBP_Cortex02.UMR_MeDIP, V8 >= 2 & V7 <= 0.05 & V16 <= 0.05), UMR = "Cortex.UMR", FDR = -log10(V7), Term = V3), UMR, Term, FDR),
                                 select(mutate(filter(GOBP_GE02.UMR_MeDIP, V8 >= 2 & V7 <= 0.05 & V16 <= 0.05), UMR = "GE.UMR", FDR = log10(V7), Term = V3), UMR, Term, FDR))
Terms <- unique(c(arrange(filter(GOBP_Cortex02_GE02_MeDIP, UMR == "Cortex.UMR"), desc(FDR))[1:20, "Term"], rev(arrange(filter(GOBP_Cortex02_GE02_MeDIP, UMR == "GE.UMR"), FDR)[1:20, "Term"])))
GOBP_Cortex02_GE02_MeDIP <- mutate(filter(GOBP_Cortex02_GE02_MeDIP, Term %in% Terms), Term = factor(Term, levels = rev(Terms)))
(GOBP_Cortex02_GE02_MeDIP_figure <- ggplot(data = GOBP_Cortex02_GE02_MeDIP, aes(Term, FDR, fill = UMR)) +
  geom_bar(position = "identity", stat = "identity", width = .5) + 
  coord_flip() + 
  xlab("") + 
  ylab("-log10(FDR)") + 
  scale_fill_manual(values = c("blue", "red"), name = "") + 
  theme_bw())
ggsave(GOBP_Cortex02_GE02_MeDIP_figure, file = "~/快盘/FetalBrain/Figures/FigureS9.pdf")
```

### Supplemental 10: DAVID enrichment for neurospheres DE genes

```{r neurospheres_DE_DAVID}
DAVID_cortex_GE_UP <- read.delim("~/快盘/FetalBrain/RNAseq/DEfine/gene/enrich/cortex_GE_UP_duplicated_DAVID.txt", as.is = T) %>% filter(Category == "GOTERM_BP_5")
DAVID_cortex_GE_DN <- read.delim("~/快盘/FetalBrain/RNAseq/DEfine/gene/enrich/cortex_GE_DN_duplicated_DAVID.txt", as.is = T) %>% filter(Category == "GOTERM_BP_5")
Terms <- unique(c(DAVID_cortex_GE_UP[DAVID_cortex_GE_UP$FDR < 0.01, "Term"], DAVID_cortex_GE_DN[DAVID_cortex_GE_DN$FDR < 0.01, "Term"]))
DAVID_neurospheres <- rbind(filter(DAVID_cortex_GE_UP, Term %in% Terms) %>% mutate(DE = "UP"), filter(DAVID_cortex_GE_DN, Term %in% Terms) %>% mutate(DE = "DN")) %>% mutate(Term = gsub("GO:\\d+~", "", Term), FDR = -log10(FDR) -2) %>% select(DE, Term, FDR)
DAVID_neurospheres$Term <- factor(DAVID_neurospheres$Term, levels = rev(unique(DAVID_neurospheres[order(DAVID_neurospheres$FDR, decreasing = T), "Term"])))
(DAVID_neurospheres_figure <- ggplot(DAVID_neurospheres, aes(x = Term, y = FDR, fill = DE)) + 
   geom_bar(stat = "identity", position = "dodge", width = 0.5) + 
   coord_flip() + 
   scale_fill_manual(values = c("blue", "red"), labels = c("UP in GE", "UP in Cortex"), name = "") + 
   scale_y_continuous(breaks = seq(-4, 6, by = 2), labels = seq(-2, 8, by = 2)) + 
   xlab("") + 
   ylab("-log10(FDR)") + 
   theme_bw())
ggsave(DAVID_neurospheres_figure, file = "~/快盘/FetalBrain/Figures/FigureS10.pdf")
```

### Supplemental 11: DAVID enrichment for neurospheres isoform genes

```{r neurospheres_isoform_DAVID}
(isoform_neurospheres_DAVID <- isoform_cortex_GE_enrich$figure + ggtitle("") + xlab(""))
ggsave(isoform_neurospheres_DAVID, file = "~/快盘/FetalBrain/Figures/FigureS11.pdf")
```

### Supplemental 12: DNA methylation at exon boundaries with MeDIP

```{r neurospheres_epiProfile_5mC_MeDIP}
(MeDIP_figure_HuFNSC01 <- MeDIP_figure_HuFNSC01 + ggtitle("HuFNSC01") + theme(legend.position = "none"))
(MeDIP_figure_HuFNSC02 <- MeDIP_figure_HuFNSC02 + ggtitle("HuFNSC02"))
pdf("~/快盘/FetalBrain/Figures/FigureS12.pdf", height = 7, width = 15)
grid.arrange(MeDIP_figure_HuFNSC01, MeDIP_figure_HuFNSC02, nrow = 1, widths = 2:3)
dev.off()
```



<!--
### Supplemental : GREAT analysis on UMRs between MZ twins - GOBP 

```{r MZ_GREAT, eval = F}
GREAT_MZ <- rbind(data.frame(UMR = "HuFNSC01.UMRs", Cell = "Brain", GREAT_HuFNSC01.UMR.Brain$data), data.frame(UMR = "HuFNSC02.UMRs", Cell = "Brain", GREAT_HuFNSC02.UMR.Brain$data), data.frame(UMR = "HuFNSC01.UMRs", Cell = "Cortex", GREAT_HuFNSC01.UMR.Cortex$data), data.frame(UMR = "HuFNSC02.UMRs", Cell = "Cortex", GREAT_HuFNSC02.UMR.Cortex$data), data.frame(UMR = "HuFNSC01.UMRs", Cell = "GE", GREAT_HuFNSC01.UMR.GE$data), data.frame(UMR = "HuFNSC02.UMRs", Cell = "GE", GREAT_HuFNSC02.UMR.GE$data))
GREAT_MZ <- filter(GREAT_MZ, Category == "GOBP")
GREAT_MZ$FDR <- -log10(GREAT_MZ$FDR)
GREAT_MZ[GREAT_MZ$UMR == "HuFNSC01.UMRs", "FDR"] <- -GREAT_MZ[GREAT_MZ$UMR == "HuFNSC01.UMRs", "FDR"]
(GREAT_MZ_figure <- ggplot(data = GREAT_MZ, aes(Term, FDR)) +
  geom_bar(aes(fill = UMR), stat = "identity", width = .5, position = "identity") + 
  coord_flip() + 
  facet_grid(Cell ~ ., scales = "free_x") + 
  xlab("") + 
  theme_bw())
ggsave(GREAT_MZ_figure, file = "~/快盘/FetalBrain/Figures/FigureS.pdf", height = height, width = width)
```
-->




