---
title: "FetalBrain - Hydroxymethylation"
author: "Gloria Li"
date: "November 4, 2014"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_depth: 4
---

Updated: `r date()`

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(xtable)
library(VennDiagram)
load("~/快盘/FetalBrain/WGBS_MeDIP/hydroxy.Rdata")
knitr::opts_chunk$set(message=FALSE, echo = FALSE, warning = FALSE)
```

## Rationale

  * WGBS: 5mc + 5hmC
  * MeDIP: 5mC
  * 5hmC: high WGBS fractional call & low MeDIP fractional call
  * Samples with both WGBS and MeDIP data: Cortex - HuFNSC02 and GE - HuFNSC02
  
## Distribution of WGBS - MeDIP

  * The differences between WGBS and MeDIP fractional calls are asymmetry, with more high WGBS low MeDIP CpGs. _Any technical bias that may contribute to this?_     
  * Potential hydroxymethylated CpGs: WGBS-MeDIP >= 0.8.    
  
```{r distribution}
(WGBS_MeDIP_diff_summary_figure + ggtitle("Differences between WGBS and MeDIP fractional calls"))
```

## Potential ydroxymethylated CpGs 

  + On average, `r round(mean(hydroxy.summary$CpG)/28217448*100, 2)`% CpGs are potentially hydroxymethylated, similar to [previous reports](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3009720/). `r round(Cortex_GE_hydroxy_CpG/mean(hydroxy.summary$CpG)*100, 2)`% of them are shared by two cell types, highly significant (hypergeometric test p = 0).    
  + On average, around 5% WGBS UMRs have potential hydroxymethylated CpGs.   
  
```{r summary, results='asis'}
hydroxy.summary <- xtable(hydroxy.summary)
align(hydroxy.summary) <- "ccccc"
print(hydroxy.summary, type = "html", include.rownames = T)

plot.new()
grid.text("Intersect of Cortex and GE potential hydroxymethylation CpGs", x = unit(0.5, "npc"), y = unit(0.95, "npc"))
grid.draw(Venn_Cortex_GE_hydroxy_CpG)
```

## Genomic breakdown 

  * Potential hydroxymethylated CpGs are mostly in introns and intergenetic regions (non-coding regions), __opposite__ from [previous reports](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3516134/). 
  
```{r genomicBreak}
(genomicBreak_hydroxy_figure + ggtitle("Genomic breakdown of potential hydroxy CpGs in Cortex02 and GE02"))
```

## GREAT analysis

  * GREAT functional enrichment on potential hydroxymethylated regions show similar terms between Cortex and GE, but not very specific to brain development.   
  
```{r GREAT, fig.height=10, fig.width=8}
(GREAT_Cortex02.hydroxy)
(GREAT_GE02.hydroxy)
```
